
Assignment : This tells the way assignments can be made in R
```{r}
  5->a
  a
  a<- 5
  a
  a=5
  a
```
Vectors: Some basic vector operations

```{r}
    vec1 <- c(1,2,3,4,5)
   class(vec1)

   vec2<-c(TRUE,FALSE)
   vec2
   class(vec2)
   vec3<-c('A','B')
   vec3
   class(vec3)
   ```

   cannot mix data types if mixed r will try to coerce into another data type

```{r}
vec4<-c(FALSE,2,3)
 vec4
```

Vec operation and some basic builtin functions

```{r}
    vec1*10
   heights=c(150,160,170)
   heights
   names(heights) <- c('alice','bob','charlie')
   heights
   heights['bob']
   sum(heights)
   mean(heights)
   max(heights)
   min(heights)
   heights[1]
   heights[c(1,2)]
``` 


another way to reference vectors


```{r}
newv <- c(1,2,3,4,5,6,7,8,9)
newv[1:4]
newv[2:4]
v1 <- 1:10
v1
```

boolean masking
================
``` {r}
    newv>2
    newv[newv>2]
```



 

matrix
============
  
```{r}   
   linkedin <- c(10,20,33,44,55)
   facebook <- c(100,200,330,440,550)
   matsocial <- c(linkedin,facebook)
   matviews<- matrix(matsocial,byrow=T,nrow=2)
   matviews
   matviews * matviews
   matviews %*% t(matviews)
   colSums(matviews)
```
similary try using rowSums,rowMeans,rbind,cbind

quantitative, qualitative data (categorical)

nominal,
ordinal

```{r}
   classes <- c("lkg","ukg","first","lkg")
   fac.classes<-factor(classes,ordered=T,levels=c("lkg","ukg","first"))
   fac.classes
   summary(fac.classes)
```
   try to use cbind and rbind to rowbind and column bind matrix


```{r}
 mat <- matrix(1:25,byrow=T,nrow=5)
 class(mat)
 is.matrix(mat)
 mat[2:3,2:3]
 a <- c(1,2,3)
 b <- c(4,5,6)
 rbind(a,b)
 class(rbind(a,b))
```
Simple plotting histogram use ?hist to get to see function definition

```{r}
hist((rnorm(20,1,20)))
hist((runif(20,1,20)))
```
************************data frames*************************
  
``` {r} 
class(ToothGrowth)
     summary(ToothGrowth)

     str(ToothGrowth)
```
******sapply,lapply,tapply*******
  
```{r}
    sapply(1:5,function(x) x^2)
    a <-lapply(1:3, function(x) x^2)
    a[1]
    unlist(a)
    tapply(mtcars$mpg,mtcars$cyl,mean)
```