---
title: "rprogbasics-contd"
author: "David Raj"
date: "20 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Install and load the necessary packages if packages are not present already install using install.packages("packagename")
```{r load packages}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(nycflights13)
```
dplyr basic funtions

SELECT,FILTER,ARRANGE,PIPELINING(not command but redirecting using %>%),MUTATE,SUMMARIZE
```{r dplyr commands}
select(mtcars,mpg:hp)

select(mtcars,contains("m"))

select(mtcars,starts_with("m"))

select(mtcars,ends_with("m"))

select(flights,-(year:day))

rename(flights,tail_num=tailnum)

select(flights,distance,everything())

filter(mtcars,cyl==6)

filter(mtcars,cyl==6,hp==110)

filter(mtcars,cyl %in% c(4,6))

mtcars %>% select(mpg,cyl,disp) %>% filter(cyl==6)

arrange(mtcars,desc(cyl,mpg))

mutate(mtcars,newhp=hp+10)
head(mtcars)
mtcars=mutate(mtcars,newhp=hp)

transmute(flights,gain=arr_delay-dep_delay)

mtcars %>% group_by(cyl) %>% summarise(avg_mpg=mean(mpg))


```

GGPLOT2 basics
This document will discuss 1) Scatter plot
                           2) Bar plot
                           3) lineplot
                           4) boxplot
                           
There are different ways of plotting the best one is the grammer of graphics using ggplot2 , in addition
there are other ways like base plotting ,lattice plotting and quick way using qplot


```{r ggplot commands}
ggplot(data=mtcars,aes(x=wt,y=mpg,col=cyl)) + geom_point()
ggplot(data=mtcars,aes(x=wt,y=mpg,col=cyl)) + geom_point(size=5)
ggplot(data=mtcars,aes(x=wt,y=mpg,col=factor(cyl))) + geom_point(size=5)

ggplot(data=mtcars,aes(x=wt,y=mpg,col=factor(hp))) + geom_point(size=5)

ggplot(data=mtcars,aes(x=wt,y=mpg,col=hp)) + geom_point(size=5)

ggplot(data=mtcars,aes(x=wt,y=mpg,col=hp)) + geom_point(size=5)+scale_color_gradient(low="blue",high="red")

ggplot(data=mtcars,aes(x=wt,y=mpg,col=factor(cyl))) + geom_point(size=5)+facet_grid(cyl~.)

ggplot(data=iris,aes(x=Sepal.Length,y=Petal.Length,col=Species))+ geom_point(size=5)

ggplot(data=iris,aes(x=Sepal.Length))+geom_histogram()

ggplot(data=iris,aes(x=Sepal.Length))+geom_histogram(binwidth=0.1)

ggplot(data=iris,aes(x=Sepal.Length,fill=Species))+geom_histogram(binwidth=0.2)

ggplot(data=iris,aes(x=Sepal.Length,fill=Species))+geom_histogram(binwidth=0.2,position="dodge")

ggplot(data=iris,aes(x=Sepal.Length,fill=Species))+geom_histogram(binwidth=0.2,position="fill")

posn_d=position_dodge(0.2)

ggplot(data=mtcars,aes(x=cyl,fill=am))+geom_bar(position=posn_d)

ggplot(data=mtcars,aes(x=cyl,fill=am))+geom_bar(position=posn_d,alpha=0.6)

ggplot(mtcars, aes(mpg, col = cyl)) +  geom_freqpoly(binwidth = 1)

head(economics)

ggplot(economics, aes(x = date, y = unemploy)) +geom_line()
ggplot(economics, aes(x = date, y = unemploy/pop)) +geom_line()

sal= c(1,2,5,6,5,4,10)

boxplot(sal)

```
In the scatter plot 1 though the cylinder variable takes values 4,6 and 8 the values are shown as continous
in the legend this is corrected in the subsequent plots by applying the factor of the variable
bar plot is generally used to get the frequency of occurence or data distribution.
Line plot is mostly used for time series data. Box plot is another simple way of getting the data distribution also 
getting the outliers. review the data and the plots and they are self explanatory

TIDY DATA: using some of the tidy verse package commands we can clean the data to get it into popular tabular format.
GATHER is database equivalent of unpivot
SPREAD is database equivalent of pivot
SEPERATE to split
UNITE to concatenate


```{r tidyverse commands}
table2
table3
table4a
table5
table4a %>% gather(`1999`,`2000`,key="year",value="population")
table2 %>% spread(key=type,value=count)
table3 %>% separate(rate,into=c("cases","population"),sep="/")
table5 %>% unite(nyear,century,year,sep="")

```

Joins using dplyr - Lets create some dummy tables and issue inner and outer joins.

```{r dplyr join commands}
x = tribble(~key,~val_x,1,"x1",2,"x2",3,"x3")
y = tribble(~key,~val_y,1,"y1",2,"y2",4,"y4")
x %>% left_join(y,by="key")
x %>% right_join(y,by="key")
x %>% full_join(y,by="key")
x %>% inner_join(y,by="key")
flights2 <- flights %>% select(-origin,-dest) %>% left_join(airlines,by="carrier")
topdest <- flights %>% count(dest,sort=TRUE) %>% head(10)
topdestfullname <- topdest %>% left_join(airports,by=c("dest"="faa"))
flights %>% semi_join(topdest)
flights %>% anti_join(planes,by="tailnum") %>% count(tailnum,sort=TRUE)
```

These commands are self explanatory. This concludes basics of dplyr,ggplot2,tidyverse 3 crucial packages in
R in addition to the base package. This will help to do lot of cleaning, transformation and visualization of data.









